---
title: "Google Data Analytics Capstone: Case Study 1"
output: html_document
date: "2023-07-24"
---

# Introduction

In this case study, we will perform data analysis for a fictional bike-share company in order to help them attract more riders. Along the way, we will perform numerous real-world tasks of a junior data analyst by following the steps of the data analysis process: **Ask, Prepare, Process, Analyze, Share,** and **Act**.

## Scenario

You are a junior data analyst working in the marketing analyst team at Cyclistic, a bike-share company in Chicago. The director of marketing believes the company's future success depends on maximizing the number of annual memberships. Therefore, your team wants to understand how casual riders and annual members use Cyclistic bikes differently. From these insights, your team will design a new marketing strategy to convert casual riders into annual members. But first, Cyclistic executives must approve your recommendations, so they must be backed up with compelling data insights and professional data visualizations.

## Characters and Teams

● **Cyclistic**: A bike-share program that features more than 5,800 bicycles and 600 docking stations. Cyclistic sets itself apart by also offering reclining bikes, hand tricycles, and cargo bikes, making bike-share more inclusive to people with disabilities and riders who can't use a standard two-wheeled bike. The majority of riders opt for traditional bikes; about 8% of riders use the assistive options. Cyclistic users are more likely to ride for leisure, but about 30% use them to commute to work each day.

● **Lily Moreno**: The director of marketing and your manager. Moreno is responsible for the development of campaigns and initiatives to promote the bike-share program. These may include email, social media, and other channels.

● **Cyclistic marketing analytics team**: A team of data analysts who are responsible for collecting, analyzing, and reporting data that helps guide Cyclistic marketing strategy. You joined this team six months ago and have been busy learning about Cyclistic's mission and business goals --- as well as how you, as a junior data analyst, can help Cyclistic achieve them.

● **Cyclistic executive team**: The notoriously detail-oriented executive team will decide whether to approve the recommended marketing program.

# Ask

● Three questions will guide the future marketing program: 1. How do annual members and casual riders use Cyclistic bikes differently? 2. Why would casual riders buy Cyclistic annual memberships? 3. How can Cyclistic use digital media to influence casual riders to become members?

Moreno has assigned you the **first question** to answer: How do annual members and casual riders use Cyclistic bikes differently?

## Guiding Questions:

**● What is the problem you are trying to solve?**

Understanding how casual riders and annual members use Cyclistic bikes differently to maximize the number of annual memberships.

**● How can your insights drive business decisions?**

It can help design a new marketing strategy to convert casual riders into annual members.

## Key tasks

**● Identify the business task**

**● Consider key stakeholders**

## Deliverable

**● A clear statement of the business task**

# Prepare

we will use Cyclistic's historical trip data to analyze and identify trends. You can download the data from this [link](https://divvy-tripdata.s3.amazonaws.com/index.html) For the purposes of this case study, the datasets are appropriate and will enable you to answer the business questions. The data has been made available by Motivate International Inc. under [this](https://ride.divvybikes.com/data-license-agreement) license.)

## Guiding questions

**• Where is your data located?**

The data is located in the link above

**• How is the data organized?**

The data is organized by month

**• Are there issues with bias or credibility in this data? Does your data ROCCC?**

No, there aren't any issues with bias or credibility in this data because it is primary data that is collected by divvybikes. Yes, the data ROCC's (reliable, organized, comprehensive, current and cited)

**• How are you addressing licensing, privacy, security, and accessibility?**

The data is public and riders' personally identifiable information is prohibited from being used, as stated by the license above.

**• How did you verify the data's integrity?**

The data integrity is verified by using RStudio to check the overall accuracy, completeness, and consistency of data.

**• How does it help you answer your question?**

The data will give us insights to how members and casual riders differ.

**• Are there any problems with the data?**

Yes, there are some problems with the data, missing values and duplicate rows.

## Key tasks:

**1. Download data and store it appropriately**

**2. Identify how it's organised**

**3. Sort and filter the data.**

**4. Determine the credibility of the data.**

## Deliverable

**● A description of all data sources used**

# Process

The 12 csv files will be combined and new columns will be added for further analysis.

## Guiding questions

-   **What tools are you choosing and why?**

Rstudio, because it can be used to clean, analyze, and visualize data.

**● Have you ensured your data's integrity?**

Yes, The data accuracy, completeness, and consistency is ensured.

**● What steps have you taken to ensure that your data is clean?**

The data is cleaned by removing missing values and duplicate rows.

**● How can you verify that your data is clean and ready to analyze?**

Checking for missing values and duplicate rows from the data and making sure each column has the right class.

**● Have you documented your cleaning process so you can review and share those results?**

Yes, the cleaning process has been documented in this file

## Key tasks

**1. Check the data for errors.**

**2. Choose your tools.**

**3. Transform the data so you can work with it effectively.**

**4. Document the cleaning process.**

## Deliverable

**● Documentation of any cleaning or manipulation of data**

## Code

### Import data and install packages

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(ggplot2)

jul_22<-read.csv("202207-divvy-tripdata.csv")
aug_22<-read.csv("202208-divvy-tripdata.csv")
sep_22<-read.csv("202209-divvy-publictripdata.csv")
oct_22<-read.csv("202210-divvy-tripdata.csv")
nov_22<-read.csv("202211-divvy-tripdata.csv")
dec_22<-read.csv("202212-divvy-tripdata.csv")
jan_23<-read.csv("202301-divvy-tripdata.csv")
feb_23<-read.csv("202302-divvy-tripdata.csv")
mar_23<-read.csv("202303-divvy-tripdata.csv")
apr_23<-read.csv("202304-divvy-tripdata.csv")
may_23<-read.csv("202305-divvy-tripdata.csv")
jun_23<-read.csv("202306-divvy-tripdata.csv")
```

After importing the dataset, we will combine the dataset using the rbind() function.

```{r}
bikeshare_df <- rbind(jul_22,aug_22,sep_22,oct_22,nov_22,dec_22,jan_23,feb_23,mar_23,apr_23,may_23,jun_23)
str(bikeshare_df)
```

After checking the data, we can see that the variable class for column "started_at" and "ended_at" is character so we will change it to POSIXlt. This will make converting times easier. We will also add The "day_of_week" and "month" columns, then sorting them for analysis purposes.

```{r}
bikeshare_df$started_at<-as.POSIXlt(bikeshare_df$started_at)
bikeshare_df$ended_at<-as.POSIXlt(bikeshare_df$ended_at)
str(bikeshare_df)
```

```{r}
bikeshare_df2 <- bikeshare_df %>%
  add_column(ride_length= difftime(.$ended_at, .$started_at, units = "mins") , .before =  "start_station_name") %>%
  add_column(day_of_week = weekdays(bikeshare_df$started_at), .after = "ride_length") %>% 
  add_column(month = months(bikeshare_df$started_at), .after = "day_of_week")
bikeshare_df2$day_of_week <- ordered(bikeshare_df2$day_of_week, levels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
bikeshare_df2$month <- ordered(bikeshare_df2$month, levels=c("July", "August", "September", "October", "November", "December", "January", "Feebruary", "March", "April", "May", "June"))

str(bikeshare_df2)
```

### Data Cleaning

The next step is to clean the data. We will use the clean-names() function so all column names are unique and consist only of the \_ character, numbers, and letters. We will then remove Na values, Null values, duplicates, and filter out trips with ride length less than 0.

```{r}
bikeshare_df2 <- bikeshare_df2 %>%
  clean_names() %>%
  drop_na() %>%
  filter(.$ride_length > 0) %>%
  distinct()

bikeshare_df2 <- bikeshare_df2[!(bikeshare_df2$start_station_name=="" | bikeshare_df2$start_station_id=="" | bikeshare_df2$end_station_name=="" | bikeshare_df2$end_station_id==""),]

```

# Analyze

## Guiding Questions

**● How should you organize your data to perform analysis on it?**

Making sure the data is cleaned and properly formatted

**● Has your data been properly formatted?**

Yes, it has been properly formatted

**● What surprises did you discover in the data?**

There are many observations that contain missing values and invalid data.

**● How will these insights help answer your business questions?**

The insights will help drive business decisions.

## Key tasks

**1. Aggregate your data so it's useful and accessible.**

**2. Organize and format your data.**

**3. Perform calculations.**

**4. Identify trends and relationships.**

## Deliverable

**● A summary of your analysis**

## Code

```{r}
#Ride length average between customer types
aggregate(bikeshare_df2$ride_length ~ bikeshare_df2$member_casual, FUN = mean)

```

```{r}
#Ride length average between bike types
aggregate(bikeshare_df2$ride_length ~ bikeshare_df2$rideable_type, FUN = mean)
```

```{r}
#bike types

bikeshare_df2 %>%
  group_by(member_casual, rideable_type) %>%
  summarise(ride_total = n(), average_ride_length = mean(ride_length)) %>%
  arrange(member_casual, rideable_type)
```

```{r}
#Ride length by month

bikeshare_df2 %>%
  group_by(member_casual, month) %>%
  summarise(ride_total = n(), average_ride_length = mean(ride_length)) %>%
  arrange(member_casual, month)
```

```{r}
#ride length by day

bikeshare_df2 %>%
  group_by(member_casual, day_of_week) %>%
  summarise(ride_total = n(), average_ride_length = mean(ride_length)) %>%
  arrange(member_casual, day_of_week)
```

```{r}
#start stations
bikeshare_df2 %>%
  group_by(member_casual, start_station_name) %>%
  summarise(ride_total = n(), average_ride_length = mean(ride_length)) %>%
  arrange(member_casual, start_station_name) %>%
  arrange(desc(ride_total))
```

```{r}
bikeshare_df2 %>%
  group_by(member_casual, end_station_name) %>%
  summarise(ride_total = n(), average_ride_length = mean(ride_length)) %>%
  arrange(member_casual, end_station_name) %>%
  arrange(desc(ride_total))
```

# Share

## Guiding questions

**● Were you able to answer the question of how annual members and casual riders use Cyclistic bikes differently?**

Yes

**● What story does your data tell?**

The data tells the difference of members and casual riders' behaviors.

**● How do your findings relate to your original question?**

My findings shows how differently members use Cyclistic bikes, which relates to the original question.

**● Who is your audience? What is the best way to communicate with them?**

My audience will be the Cyclistic executive team.

**● Can data visualization help you share your findings?**

Yes, by visualizing the data, we will be able to communicate our findings to our audience.

**● Is your presentation accessible to your audience?**

The presentation will be accessible at [This link](https://github.com/RifqiHafizuddin/Google-Data-Analytics-Capstone-Bike-Share)

## Key tasks

**1. Determine the best way to share your findings.**

**2. Create effective data visualizations.**

**3. Present your findings.**

**4. Ensure your work is accessible.**

## Deliverable

**● Supporting visualizations and key findings**

## Visualization

```{r}
#member vs casual customers

bikeshare_df2 %>%
  group_by(member_casual) %>%
  summarise(user_total = n()) %>%
  arrange(member_casual) %>%
  ggplot(aes(x=member_casual, y=user_total, fill = member_casual)) +
  geom_col() +
  geom_text(aes(label = user_total), position = position_dodge(0.9), vjust = 0) +
  labs(title = "Members vs Casual")
```

```{r}
#Weekly number of rides by riders
bikeshare_df2 %>%
  group_by(member_casual, day_of_week) %>%
  summarise(ride_total = n()) %>%
  arrange(member_casual, day_of_week) %>%
  ggplot(aes(x = day_of_week, y = ride_total, fill = member_casual)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = ride_total), position = position_dodge(0.9), vjust = 0, size = 3) +
  labs(title = "Weekly number of rides by riders")
```

```{r}
#monthly number of rides by riders
bikeshare_df2 %>%
  group_by(member_casual, month) %>%
  summarise(ride_total = n()) %>%
  arrange(member_casual, month) %>%
  ggplot(aes(x = month, y = ride_total, fill = member_casual)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = ride_total), position = position_dodge(0.9), vjust = 0, size = 3) +
  labs(title = "Monthly Number of Rides by Riders")
```

```{r}
#Weekly average ride length by riders
bikeshare_df2 %>%
  group_by(member_casual, day_of_week) %>%
  summarise(average_ride_length = mean(ride_length)) %>%
  arrange(member_casual, day_of_week) %>%
  ggplot(aes(x = day_of_week, y = average_ride_length, fill = member_casual)) +
  geom_col(position = "dodge") +
  labs(title = "Weekly average ride length by riders")
```

```{r}
bikeshare_df2 %>%
  group_by(member_casual, month) %>%
  summarise(average_ride_length = mean(ride_length)) %>%
  arrange(member_casual, month) %>%
  ggplot(aes(x = month, y = average_ride_length, fill = member_casual)) +
  geom_col(position = "dodge") +
  labs(title = "monthly average ride length by riders")
```

## Summary

-   Casual riders rode the bike longer (22.59 mins) than members (12.17 mins).

-   The average ride length of the docked bike (49.81 mins) is longer than electric bike (12.7 mins) and classic bike (16.78 mins).

-   The most popular start and end station for the casual riders is Streeter Dr & Grand Ave, while the members start and end their rides at Kingsbury St & Kinzie St.

-   There are more members (2616490) than casual riders (1642708) in the last 12 months

-   On average, casual riders ride the bike more on weekends, while members ride the bike more on weekdays.

-   Both members and casual riders ride the bike more near summer and ride it less near winter.

-   The average ride length of casual riders are longer than members.

# Act

## Guiding questions

**● What is your final conclusion based on your analysis?**

Based on my analysis, we can see that on average, casual riders ride the bike longer than members, so converting casual rider into annual members is possible.

**● How could your team and business apply your insights?**

To emphasize on the casual riders' longer ride time, we could develop a marketing campaign that focuses on the benefits of being a member while riding the bike longer.

**● What next steps would you or your stakeholders take based on your findings?**

The next steps we would take is to apply our insights and if possible, use additional data to expand on our findings.

**● Is there additional data you could use to expand on your findings?**

Yes, additional data such as user demographics, financial data, past user purchases (annual vs single pass), and climate data.

## Key tasks

**1. Create your portfolio.**

**2. Add your case study.**

**3. Practice presenting your case study to a friend or family member.**

## Deliverable

**● Your top three recommendations based on your analysis.**

1.  Marketing campaign that targets casual users who have high average ride times and buys many single passes.

2.  Promote the campaign at popular start stations for casual riders such as Streeter Dr & Grand Ave, DuSable Lake Shore Dr & Monroe St, or Michigan Ave & Oak St.

3.  Give discounts for annual passes at less popular months to increase members.
